Chord: A Versatile Platform for Program Analysis

### 2011-03-14 main/doc/dynamic_analysis.tex
more documentation

@@

\subsection{Implementing the Analysis}

@@

\subsection{Configuring the Analysis}

You can change the default values of various properties for configuring your dynamic analysis;
see ``Analysis Scope Properties" and ``Instrumentation Properties" in Section \ref{sec:chord-sysprops}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous Chord run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

\subsection{Compiling and Running the Analysis}

@@

>>>>

\chapter{Dynamic Analysis}
\label{chap:dynamic-analysis}

This chapter describes everything about dynamic analysis in Chord.
Section \ref{sec:writing-dynamic-analysis} describes how to write a dynamic analysis,
Section \ref{sec:running-dynamic-analysis} describes how to compile and run it,
and Section \ref{sec:instr-events} describes common dynamic analysis events
supported in Chord.

@@

@@

\section{Compiling and Running a Dynamic Analysis}
\label{sec:running-dynamic-analysis}

@@

You can change the default values of various properties for configuring your dynamic analysis;
see Section \ref{sec:scope-props} and Section \ref{sec:instr-props} in Chapter \ref{chap:properties}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous Chord run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

\section{Common Dynamic Analysis Events}
\label{sec:instr-events}

Chord provides support for instrumenting common dynamic analysis events.
The below table describes these events.

\begin{table}
\begin{center}
\htmlattributes*{table}{border="1" cellpadding="10"}
\begin{tabular}{|l|p{4.3in}|} 
\hline
{\bf Event Kind} & {\bf Description} \\
\hline
EnterMainMethod(\bt) & After thread \bt\ enters method \bm\ (in domain M).
\\
\hline
EnterMethod(\bm, \bt) & After thread \bt\ enters method \bm\ (in domain M).
\\
\hline
LeaveMethod(\bm, \bt) & Before thread \bt\ leaves method \bm\ (in domain M).
\\
\hline
EnterLoop(\bw, \bt) & Before thread \bt\ begins loop \bw\ (in domain W).
\\
\hline
LoopIteration(\bw, \bt) & Before thread \bt\ starts a new iteration of loop \bw\ (in domain W).
\\
\hline
LeaveLoop(\bw, \bt) & After thread \bt\ finishes loop \bw\ (in domain W).
\\
\hline
BasicBlock(\bb, \bt) & Before thread \bt\ enters basic block \bb\ (in domain B).
\\
\hline
Quad(\bp, \bt) & Before thread \bt\ executes quad at program point \bp\ (in domain P).
\\
\hline
BefMethodCall(\bi, \bt, \bo) & 
Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
{\bf Note:} This event is not generated before constructor calls; use the BefNew event.
\\
\hline
AftMethodCall(\bi, \bt, \bo) &
After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
{\bf Note:} This event is not generated after constructor calls; use the AftNew event.
\\
\hline
BefNew(\bh, \bt, \bo) & Before thread \bt\ executes a \code{new} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
\\
\hline
AftNew(\bh, \bt, \bo) & After thread \bt\ executes a \code{new} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
\\
\hline
NewArray(\bh, \bt, \bo) & After thread \bt\ executes a \code{newarray} bytecode instruction and allocates fresh object \bo\ at program point \bh\ (in domain H).
\\
\hline
GetstaticPrimitive(\be, \bt, \bg) & After thread \bt\ reads primitive-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
\\
\hline
GetstaticReference(\be, \bt, \bg, \bo) & After thread \bt\ reads object \bo\ from reference-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
\\
\hline
PutstaticPrimitive(\be, \bt, \bg) & After thread \bt\ writes primitive-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
\\
\hline
PutstaticReference(\be, \bt, \bg, \bo) & After thread \bt\ writes object \bo\ to reference-typed static field \bg\ (in domain F) at program point \be\ (in domain E).
\\
\hline
GetfieldPrimitive(\be, \bt, \bb, \bg) & After thread \bt\ reads primitive-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
\\
\hline
GetfieldReference(\be, \bt, \bb, \bg, \bo) & After thread \bt\ reads object \bo\ from reference-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
\\
\hline
PutfieldPrimitive(\be, \bt, \bb, \bg) & After thread \bt\ writes primitive-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
\\
\hline
PutfieldReference(\be, \bt, \bb, \bg, \bo) & After thread \bt\ writes object \bo\ to reference-typed instance field \bg\ (in domain F) of object \bb\ at program point \be\ (in domain E).
\\
\hline
AloadPrimitive(\be, \bt, \bb, \bi) & After thread \bt\ reads the primitive-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
\\
\hline
AloadReference(\be, \bt, \bb, \bi, \bo) & After thread \bt\ reads object \bo\ from the reference-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
\\
\hline
AstorePrimitive(\be, \bt, \bb, \bi) & After thread \bt\ writes the primitive-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
\\
\hline
AstoreReference(\be, \bt, \bb, \bi, \bo) & After thread \bt\ writes object \bo\ to the reference-typed element at index \bi\ of array object \bb\ at program point \be\ (in domain E).
\\
\hline
%ReturnPrimitive(\bp, \bt) & Not yet supported.
%\\
%ReturnReference(\bp, \bt) & Not yet supported.
%\\
%ExplicitThrow(\bp, \bt, \bo) & Not yet supported.
%\\
%ImplicitThrow(\bp, \bt, \bo) & Not yet supported.
%\\
ThreadStart(\bi, \bt, \bo) & Before thread \bt\ calls the \code{start()} method of \code{java.lang.Thread} at program point \bi\ (in domain I) and spawns a thread \bo.
\\
\hline
ThreadJoin(\bi, \bt, \bo) & Before thread \bt\ calls the \code{join()} method of \code{java.lang.Thread} at program point \bi\ (in domain I) to join with thread \bo.
\\
\hline
AcquireLock(\bl, \bt, \bo) & After thread \bt\ executes a statement of the form ``monitorenter \bo'' or enters a method synchronized on \bo\ at program point \bl\ (in domain L).
\\
\hline
ReleaseLock(\br, \bt, \bo) & Before thread \bt\ executes a statement of the form ``monitorexit \bo'' or leaves a method synchronized on \bo\ at program point \br\ (in domain R).
\\
\hline
Wait(\bi, \bt, \bo) & Before thread \bt\ calls the \code{wait()} method of \code{java.lang.Object} on object \bo\ at program point \bi\ (in domain I).
\\
\hline
Notify(\bi, \bt, \bo) & Before thread \bt\ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} on object \bo\ at program point \bi\ (in domain I).
\T \\
\hline
\end{tabular}
\end{center}
\end{table}

### 2011-03-14 main/doc/getting_started.tex
more documentation

\begin{enumerate}
\item
\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
\item
\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
These files are needed only if you want to use the BuDDy BDD library for executing program
analyses written in Datalog.
\item
\code{libchord_instr_agent.so}: this file is needed only if you want to use the JVMTI-based bytecode
instrumentation agent for executing dynamic program analyses in Chord.
\end{enumerate}

@@

\begin{enumerate}
\item
Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
\item
Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
\end{enumerate}

@@

Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
specific sub-directories:
\begin{itemize}
\item \code{main/} contains Chord's source code and the jars of libraries used by Chord.
\item \code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
\item \code{test/} contains Chord's regression tests.
\item many more; these might eventually move into \code{main/}.
\end{itemize}

@@

This will compile Chord's Java sources from \code{main/src/} to class files
in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
these class files as well as the class files in the jars of libraries that are used by Chord
and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
\code{bddbddb.jar}, etc.).
Additionally:
\begin{itemize}
\item
If system property \code{chord.use.buddy} is set to \code{true},
then the C source code of the \xlink{BuDDy}{http://buddy.sourceforge.net/} BDD library
from \code{main/bdd/} will be compiled to a shared library in \code{main/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is used for executing
program analyses written in Datalog
using \xlink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
If system property \code{chord.use.jvmti} is set to \code{true},
then the C++ source code of the JVMTI-based bytecode instrumentation agent from
\code{main/agent/} will be compiled to a shared library in \code{main/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
used for computing program scope dynamically and for executing
dynamic program analyses.
\end{itemize}
Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
default value of both these properties is \code{false}.
If either of them is set to \code{true}, then you will also need a utility like GNU Make (to run the
\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.

@@

Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to value
\code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
command-line argument processing.
Chapter\ref{chap:properties} describes all system properties recognized by Chord.

>>>>

\begin{enumerate}
\item
\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
\item
\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
These files are needed only if you want BDD library BuDDy to be used when 
the BDD-based Datalog solver bddbddb in Chord runs analyses written in Datalog.
\item
\code{libchord_instr_agent.so}: this file is needed only if you want the JVMTI-based bytecode
instrumentation agent to be used when Chord runs dynamic analyses.
\end{enumerate}

@@

\begin{itemize}
\item
Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
\item
Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
\end{itemize}

@@

Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
specific sub-directories:

\begin{itemize}
\item
\code{main/} contains Chord's source code and the jars of libraries used by Chord.
\item
\code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
\item
\code{test/} contains Chord's regression tests.
\item
many more; these might eventually move into \code{main/}.
\end{itemize}

@@

This will compile Chord's Java sources from \code{main/src/} to class files
in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
these class files as well as the class files in the jars of libraries that are used by Chord
and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
\code{bddbddb.jar}, etc.).
Additionally:

\begin{itemize}
\item

If system property \code{chord.use.buddy} is set to \code{true},
then the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} 
from \code{main/bdd/} will be compiled to a shared library in \code{main/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is used 
by BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} 
in Chord for running analyses written in Datalog.

\item

If system property \code{chord.use.jvmti} is set to \code{true},
then the C++ source code of the JVMTI-based bytecode instrumentation agent from
\code{main/agent/} will be compiled to a shared library in \code{main/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
used in Chord for computing analysis scope dynamically and for running
dynamic analyses.
\end{itemize}

Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
default value of both these properties is \code{false}.
If you set either of them to \code{true}, then you will also need a utility like GNU Make (to run the
\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.

@@

Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to
the value denoted by \code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
command-line argument processing.
Chapter\ref{chap:properties} describes all system properties recognized by Chord.
